---
title: "Preliminary Data Analysis"
author: "Olgiati Stefano"
date: "25 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(data.table)

```

# Reading tidy data

```{r echo=FALSE}


data.EU.1990 <- read.csv("Output_EU_1990.csv")
data.EU.2010 <- read.csv("Output_EU_2010.csv")
data.ROW.1990 <- read.csv("Output_ROW_1990.csv")
data.ROW.2010 <- read.csv("Output_ROW_2010.csv")

head(data.EU.1990)
head(data.EU.2010)
head(data.ROW.1990)
head(data.ROW.2010)

attach(data.EU.1990)
attach(data.EU.2010)
attach(data.ROW.1990)
attach(data.ROW.2010)

df.EU.1990 <- data.frame(country=data.EU.1990$country, le=data.EU.1990$le, hale=data.EU.1990$hale, yld=data.EU.1990$yld, the=data.EU.1990$the, gdp=data.EU.1990$gdp, the.gdp=data.EU.1990$the.gdp)

df.EU.2010 <- data.frame(country=data.EU.2010$country, le=data.EU.2010$le,hale=data.EU.2010$hale, yld=data.EU.2010$yld, the=data.EU.2010$the, gdp=data.EU.2010$gdp, the.gdp=data.EU.2010$the.gdp)

df.ROW.1990 <- data.frame(country=data.ROW.1990$country, le=data.ROW.1990$le,hale=data.ROW.1990$hale, yld=data.ROW.1990$yld, the=data.ROW.1990$the, gdp=data.ROW.1990$gdp, the.gdp=data.ROW.1990$the.gdp)

df.ROW.2010 <- data.frame(country=data.ROW.2010$country, le=data.ROW.2010$le,hale=data.ROW.2010$hale, yld=data.ROW.2010$yld, the=data.ROW.2010$the, gdp=data.ROW.2010$gdp, the.gdp=data.ROW.2010$the.gdp)

df.EU.1990
df.EU.2010
df.ROW.1990
df.ROW.2010

ID.ROW <- "JPN"
ID.EU <- "ITA"

# Constant Scale in Graphs

lim.le <- c(30, 85)
lim.yld <- c(4, 15)
lim.hale <- c(30, 85)
lim.the <- c(100, 8500)
lim.gdp <- c(1000, 87500)

```



# Expansion of Morbidity WORLD: 1990-2010

```{r}
par(mar=c(5, 4, 4, 5) + 0.1)

plot(df.ROW.1990$hale, df.ROW.1990$yld, xlab="Health Adjusted Life Expectancy at Birth (hale)", ylab="Years Lived with Disease (YLD)", type="p", col="grey", pch = 16, cex = 1, lty = "solid", lwd = 2, xlim=lim.hale, ylim=lim.yld, axes=FALSE, frame.plot=FALSE)

points(df.ROW.2010$hale, df.ROW.2010$yld, col="darkslategrey", pch = 16, cex = 1, lty = "solid", lwd = 2)

points(df.EU.1990$hale, df.EU.1990$yld, col="blue", pch = 16, cex = 1, lty = "solid", lwd = 2)

points(df.EU.2010$hale, df.EU.2010$yld,  col="red", pch = 16, cex = 1, lty = "solid", lwd = 2)

points(df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], col="black", pch = 8, cex = 1, lty = "solid", lwd = 2)

points(df.EU.1990$hale[df.EU.1990$country == ID.EU], df.EU.1990$yld[df.EU.1990$country == ID.EU],  col="green", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

points(df.EU.2010$hale[df.EU.2010$country == ID.EU], df.EU.2010$yld[df.EU.2010$country == ID.EU],  col="yellow", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

axis(1, at=c(round(min(df.EU.1990$hale), 1), round(median(df.EU.1990$hale), 1), round(max(df.EU.1990$hale), 1)), col.axis="blue", col="blue", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=round(df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black", las=2, cex.axis=0.5,  tick=-.01)

axis(2, at=c(round(min(df.EU.1990$yld), 1), round(median(df.EU.1990$yld), 1), round(max(df.EU.1990$yld), 1)), col.axis="blue", col="blue",las=2, cex.axis=0.7, tick=-.01)

axis(4, at=round(df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black", las=2, cex.axis=0.5,  tick=-.01)

axis(3, at=c(round(min(df.EU.2010$hale), 1), round(median(df.EU.2010$hale), 1), round(max(df.EU.2010$hale), 1)), col.axis="red", col="red",las=2, cex.axis=0.7,  tick=-.01)

axis(4, at=c(round(min(df.EU.2010$yld), 1), round(median(df.EU.2010$yld), 1), round(max(df.EU.2010$yld), 1)), col.axis="red", col="red", las=2, cex.axis=0.7, tick=-.01)

legend(30, 15, c("EU 1990","EU 2010","ITA 1990", "ITA 2010", "JPN 2010"), col=c("blue", "red", "green", "yellow", "black"), bty="n", y.intersp=1.5, pch=c(16,16, 9, 9, 8), cex = .75)


abline(h=df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], v=df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], lty=1, lwd=1)

fun.EU.1990 <- lm(df.EU.1990$yld ~ df.EU.1990$hale)
abline(fun.EU.1990, lty = 3, col="blue", lwd=1.5)
summary(fun.EU.1990)
confint(fun.EU.1990)

fun.EU.2010 <- lm(df.EU.2010$yld ~ df.EU.2010$hale)
abline(fun.EU.2010, lty = 3, col="red", lwd=1.5)
summary(fun.EU.2010)
confint(fun.EU.2010)
```



# Increased Expansion of Morbidity EU: 1990-2010

```{r}
par(mar=c(5, 4, 4, 5) + 0.1)

plot(df.EU.1990$hale, df.EU.1990$yld, xlab="Health Adjusted Life Expectancy at Birth (hale)", ylab="Years Lived with Disease (YLD)", col="blue", pch = 16, cex = 1.3, lty = "solid", lwd = 2, xlim=lim.hale, ylim=lim.yld, axes=FALSE, frame.plot=FALSE)

points(df.EU.2010$hale, df.EU.2010$yld,  col="red", pch = 16, cex = 1.3, lty = "solid", lwd = 2)

points(df.EU.1990$hale[df.EU.1990$country == ID.EU], df.EU.1990$yld[df.EU.1990$country == ID.EU],  col="black", pch = 5, cex = 1.3, lty = "solid", lwd = 4)

points(df.EU.2010$hale[df.EU.2010$country == ID.EU], df.EU.2010$yld[df.EU.2010$country == ID.EU],  col="black", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

axis(1, at=c(round(min(df.EU.1990$hale), 1), round(median(df.EU.1990$hale), 1), round(max(df.EU.1990$hale), 1)), col.axis="blue", col="blue", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=round(df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black", las=2, cex.axis=0.7,  tick=-.01)

axis(2, at=c(round(min(df.EU.1990$yld), 1), round(median(df.EU.1990$yld), 1), round(max(df.EU.1990$yld), 1)), col.axis="blue", col="blue",las=2, cex.axis=0.7, tick=-.01)

axis(4, at=round(df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=c(round(min(df.EU.2010$hale), 1), round(median(df.EU.2010$hale), 1), round(max(df.EU.2010$hale), 1)), col.axis="red", col="red",las=2, cex.axis=0.7,  tick=-.01)

axis(4, at=c(round(min(df.EU.2010$yld), 1), round(median(df.EU.2010$yld), 1), round(max(df.EU.2010$yld), 1)), col.axis="red", col="red", las=2, cex.axis=0.7, tick=-.01)

legend(30, 15, c("EU 1990","EU 2010","ITA 1990", "ITA 2010", "JPN 2010"), col=c("blue", "red", "black", "black", "black"), bty="n", y.intersp=1.5, pch=c(16,16, 5, 9, 8), cex = .75)

points(df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], col="black", pch = 8, cex = 1.3, lty = "solid", lwd = 2)

abline(h=df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], v=df.ROW.2010$hale[df.ROW.2010$country == ID.ROW], lty=1, lwd=1)


fun.EU.1990 <- lm(df.EU.1990$yld ~ df.EU.1990$hale)
abline(fun.EU.1990, lty = 3, col="blue", lwd=1.5)
summary(fun.EU.1990)
confint(fun.EU.1990)

fun.EU.2010 <- lm(df.EU.2010$yld ~ df.EU.2010$hale)
abline(fun.EU.2010, lty = 3, col="red", lwd=1.5)
summary(fun.EU.2010)
confint(fun.EU.2010)

```



# Growth of Health Expenditure EU: 1990-2010

```{r}
par(mar=c(5, 4, 4, 5) + 0.1)

plot(df.EU.1990$le, df.EU.1990$the, xlab="Life Expectancy at Birth", ylab="Per Capita Total Health Expenditure", col="blue", pch = 16, cex = 1.3, lty = "solid", lwd = 2, xlim=lim.le, ylim=lim.the, axes=FALSE, frame.plot=FALSE)

points(df.EU.2010$le, df.EU.2010$the,  col="red", pch = 16, cex = 1.3, lty = "solid", lwd = 2)

points(df.EU.1990$le[df.EU.1990$country == ID.EU], df.EU.1990$the[df.EU.1990$country == ID.EU],  col="black", pch = 5, cex = 1.3, lty = "solid", lwd = 4)

points(df.EU.2010$le[df.EU.2010$country == ID.EU], df.EU.2010$the[df.EU.2010$country == ID.EU],  col="black", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

legend(35, 7000, c("EU 1990","EU 2010","ITA 1990", "ITA 2010", "JPN 2010"), col=c("blue", "red", "black", "black", "black"), bty="n", y.intersp=1.5, pch=c(16,16, 5, 9, 8), cex = .75)

axis(1, at=c(round(min(df.EU.1990$le), 1), round(median(df.EU.1990$le), 1), round(max(df.EU.1990$le), 1)), col.axis="blue", col="blue", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=round(df.ROW.2010$le[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)


low.the.1990 <- round(min(df.EU.1990$the), 0)
median.the.1990 <- round(median(df.EU.1990$the), 0)
mad.the.1990 <- round(mad(df.EU.1990$the), 0)
high.the.1990 <- round(max(df.EU.1990$the), 0)

axis(2, at=c(low.the.1990,median.the.1990, high.the.1990), col.axis="blue", col="blue",las=2, cex.axis=0.7, tick=-.01)

axis(3, at=c(round(min(df.EU.2010$le), 1), round(median(df.EU.2010$le), 1), round(max(df.EU.2010$le), 1)), col.axis="red", col="red", las=2, cex.axis=0.7, tick=-.01)

low.the.2010 <- round(min(df.EU.2010$the), 0)
median.the.2010 <- round(median(df.EU.2010$the), 0)
mad.the.2010 <- round(mad(df.EU.2010$the), 0)
high.the.2010 <- round(max(df.EU.2010$the), 0)
japan.the.2010 <- round(df.ROW.2010$the[df.ROW.2010$country == ID.ROW], 1)

axis(4, at=c(low.the.2010,median.the.2010, high.the.2010), col.axis="red", col="red",las=2, cex.axis=0.7, tick=-.01)

axis(4, at=japan.the.2010, col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)

points(df.ROW.2010$le[df.ROW.2010$country == ID.ROW], df.ROW.2010$the[df.ROW.2010$country == ID.ROW], col="black", pch = 8, cex = 1.3, lty = "solid", lwd = 2)

abline(h=df.ROW.2010$the[df.ROW.2010$country == ID.ROW], v=df.ROW.2010$le[df.ROW.2010$country == ID.ROW], lty=1, lwd=1)


fun.EU.1990 <- lm(df.EU.1990$the ~ df.EU.1990$le)
abline(fun.EU.1990, lty = 2, col="blue", lwd=1.5)

fun.EU.2010 <- lm(df.EU.2010$the ~ df.EU.2010$le)
abline(fun.EU.2010, lty = 2, col="red", lwd=1.5)

summary(fun.EU.1990)
summary(fun.EU.2010)

confint(fun.EU.1990)
confint(fun.EU.2010)

```

```{r}
par(mar=c(5, 4, 4, 5) + 0.1)

plot(df.EU.1990$yld, df.EU.1990$the, xlab="Years Lived with Disability", ylab="Per Capita Total Health Expenditure", col="blue", pch = 16, cex = 1.3, lty = "solid", lwd = 2, xlim=lim.yld, ylim=lim.the, axes=FALSE, frame.plot=FALSE)

points(df.EU.2010$yld, df.EU.2010$the,  col="red", pch = 16, cex = 1.3, lty = "solid", lwd = 2)

points(df.EU.1990$yld[df.EU.1990$country == ID.EU], df.EU.1990$the[df.EU.1990$country == ID.EU],  col="black", pch = 5, cex = 1.3, lty = "solid", lwd = 4)

points(df.EU.2010$yld[df.EU.2010$country == ID.EU], df.EU.2010$the[df.EU.2010$country == ID.EU],  col="black", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

legend(5, 7000, c("EU 1990","EU 2010","ITA 1990", "ITA 2010", "JPN 2010"), col=c("blue", "red", "black", "black", "black"), bty="n", y.intersp=1.5, pch=c(16,16, 5, 9, 8), cex = .75)

axis(1, at=c(round(min(df.EU.1990$yld), 1), round(median(df.EU.1990$yld), 1), round(max(df.EU.1990$yld), 1)), col.axis="blue", col="blue", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=round(df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)


low.the.1990 <- round(min(df.EU.1990$the), 0)
median.the.1990 <- round(median(df.EU.1990$the), 0)
high.the.1990 <- round(max(df.EU.1990$the), 0)

axis(2, at=c(low.the.1990,median.the.1990, high.the.1990), col.axis="blue", col="blue",las=2, cex.axis=0.7, tick=-.01)

axis(3, at=c(round(min(df.EU.2010$yld), 1), round(median(df.EU.2010$yld), 1), round(max(df.EU.2010$yld), 1)), col.axis="red", col="red", las=2, cex.axis=0.7, tick=-.01)

low.the.2010 <- round(min(df.EU.2010$the), 0)
median.the.2010 <- round(median(df.EU.2010$the), 0)
high.the.2010 <- round(max(df.EU.2010$the), 0)
japan.the.2010 <- round(df.ROW.2010$the[df.ROW.2010$country == ID.ROW], 1)

axis(4, at=c(low.the.2010,median.the.2010, high.the.2010), col.axis="red", col="red",las=2, cex.axis=0.7, tick=-.01)

axis(4, at=japan.the.2010, col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)

points(df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], df.ROW.2010$the[df.ROW.2010$country == ID.ROW], col="black", pch = 8, cex = 1.3, lty = "solid", lwd = 2)

abline(h=df.ROW.2010$the[df.ROW.2010$country == ID.ROW], v=df.ROW.2010$yld[df.ROW.2010$country == ID.ROW], lty=1, lwd=1)


fun.EU.1990 <- lm(df.EU.1990$the ~ df.EU.1990$yld)
abline(fun.EU.1990, lty = 3, col="blue", lwd=1.5)

fun.EU.2010 <- lm(df.EU.2010$the ~ df.EU.2010$yld)
abline(fun.EU.2010, lty = 3, col="red", lwd=1.5)

summary(fun.EU.1990)
summary(fun.EU.2010)

confint(fun.EU.1990)
confint(fun.EU.2010)

```

# Expansion of Health Expenditure EU: 1990-2010

```{r}
par(mar=c(5, 4, 4, 5) + 0.1)

plot(df.EU.1990$gdp, df.EU.1990$the, xlab="Per Capita Gross Domestic Product", ylab="Per Capita Total Health Expenditure", col="blue", pch = 16, cex = 1.3, lty = "solid", lwd = 2, xlim=lim.gdp, ylim=lim.the, axes=FALSE, frame.plot=FALSE)

points(df.EU.2010$gdp, df.EU.2010$the,  col="red", pch = 16, cex = 1.3, lty = "solid", lwd = 2)

points(df.EU.1990$gdp[df.EU.1990$country == ID.EU], df.EU.1990$the[df.EU.1990$country == ID.EU],  col="black", pch = 5, cex = 1.3, lty = "solid", lwd = 4)

points(df.EU.2010$gdp[df.EU.2010$country == ID.EU], df.EU.2010$the[df.EU.2010$country == ID.EU],  col="black", pch = 9, cex = 1.3, lty = "solid", lwd = 4)

legend(1000, 7000, c("EU 1990","EU 2010","ITA 1990", "ITA 2010", "JPN 2010"), col=c("blue", "red", "black", "black", "black"), bty="n", y.intersp=1.5, pch=c(16,16, 5, 9, 8), cex = .75)

axis(1, at=c(round(min(df.EU.1990$gdp), 1), round(median(df.EU.1990$gdp), 1), round(max(df.EU.1990$gdp), 1)), col.axis="blue", col="blue", las=2, cex.axis=0.7,  tick=-.01)

axis(3, at=round(df.ROW.2010$gdp[df.ROW.2010$country == ID.ROW], 2), col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)

low.the.1990 <- round(min(df.EU.1990$the), 0)
median.the.1990 <- round(median(df.EU.1990$the), 0)
high.the.1990 <- round(max(df.EU.1990$the), 0)

axis(2, at=c(low.the.1990,median.the.1990, high.the.1990), col.axis="blue", col="blue",las=2, cex.axis=0.7, tick=-.01)

axis(3, at=c(round(min(df.EU.2010$gdp), 1), round(median(df.EU.2010$gdp), 1), round(max(df.EU.2010$gdp), 1)), col.axis="red", col="red", las=2, cex.axis=0.7, tick=-.01)

low.the.2010 <- round(min(df.EU.2010$the), 0)
median.the.2010 <- round(median(df.EU.2010$the), 0)
high.the.2010 <- round(max(df.EU.2010$the), 0)
japan.the.2010 <- round(df.ROW.2010$the[df.ROW.2010$country == ID.ROW], 1)

axis(4, at=c(low.the.2010,median.the.2010, high.the.2010), col.axis="red", col="red",las=2, cex.axis=0.7, tick=-.01)

axis(4, at=japan.the.2010, col.axis="black", col="black",las=2, cex.axis=0.7, tick=-.01)

points(df.ROW.2010$gdp[df.ROW.2010$country == ID.ROW], df.ROW.2010$the[df.ROW.2010$country == ID.ROW], col="black", pch = 8, cex = 1.3, lty = "solid", lwd = 2)

abline(h=df.ROW.2010$the[df.ROW.2010$country == ID.ROW], v=df.ROW.2010$gdp[df.ROW.2010$country == ID.ROW], lty=1, lwd=1)


fun.EU.1990 <- lm(df.EU.1990$the ~ df.EU.1990$gdp)
abline(fun.EU.1990, lty = 2, col="blue", lwd=1.5)

fun.EU.2010 <- lm(df.EU.2010$the ~ df.EU.2010$gdp)
abline(fun.EU.2010, lty = 2, col="red", lwd=1.5)

summary(fun.EU.1990)
summary(fun.EU.2010)

confint(fun.EU.1990)
confint(fun.EU.2010)

```



library(plotly)

# plot_ly(df.ROW.2010, x = hale, y = yld, text = paste("THE as % of GDP: ", country, the.gdp),  mode = "markers", color = the.gdp, size = gdp, opacity= gdp, colors = terrain.colors(3))

plot_ly(df.ROW.1990, x = hale, y = yld,  mode = "markers",  size = the, opacity= the, colors = terrain.colors(3), name="ROW 1990")

add_trace(df.ROW.2010, x = hale, y = yld, mode = "markers", size = the, opacity= the, name="ROW 2010")

plot_ly(df.EU.1990, x = hale, y = yld, mode = "markers", size = the, opacity= the, name="EU 1990")

add_trace(df.EU.2010, x = hale, y = yld, mode = "markers", size = the, opacity= the, name="EU 2010")




# Sustainable Quality EU

```{r}
df.sust.EU <- data.frame(Country=df.EU.2010$country)

df.sust.EU$DELTA_HALE <- round((df.EU.2010$hale - df.EU.1990$hale) / df.EU.1990$hale, 3) * 100

df.sust.EU$DELTA_YLD <- round((df.EU.2010$yld - df.EU.1990$yld) / df.EU.1990$yld, 3) * 100

df.sust.EU$QUALITYe <- round(df.sust.EU$DELTA_HALE / df.sust.EU$DELTA_YLD, 3)

df.sust.EU$DELTA_THE <- round((df.EU.2010$the - df.EU.1990$the) / df.EU.1990$the, 3)  * 100

df.sust.EU$DELTA_GDP <- round((df.EU.2010$gdp - df.EU.1990$gdp) / df.EU.1990$gdp, 3)  * 100

df.sust.EU$COSTe <- round(df.sust.EU$DELTA_THE / df.sust.EU$DELTA_GDP, 3)

df.sust.EU$VALUEe <- round(df.sust.EU$QUALITYe / df.sust.EU$COSTe, 3)

print(df.sust.EU, row.names = F)

plot(density(df.sust.EU$QUALITYe), main="Sustainable Universal Health Coverage", col="blue", lty=1, axes=F, lwd=1, xlab="QUALITYe, COSTe", ylab = "Frequency", ylim=c(0,3), xlim=c(0,2.5))

lines(density(df.sust.EU$COSTe), col="red", lty=1, lwd=1)

curve(dnorm(x, mean(df.sust.EU$QUALITYe), sd(df.sust.EU$QUALITYe)),
      mean(df.sust.EU$QUALITYe) - 1.96 * sd(df.sust.EU$QUALITYe),
      mean(df.sust.EU$QUALITYe) + 1.96 * sd(df.sust.EU$QUALITYe),
      add=T,col="blue", lty=2, lwd=1)

curve(dnorm(x, mean(df.sust.EU$COSTe), sd(df.sust.EU$COSTe)),
       mean(df.sust.EU$COSTe) - 1.96 * sd(df.sust.EU$COSTe),
      mean(df.sust.EU$COSTe) + 1.96 * sd(df.sust.EU$COSTe),
      add=T,col="red", lty=2, lwd=1)


axis(1)
axis(2)

abline(v=df.sust.EU$QUALITYe[df.sust.EU$Country == ID.EU], lty=1, col="blue", lwd=1.5)

abline(v=df.sust.EU$COSTe[df.sust.EU$Country == ID.EU], lty=1, col="red", lwd=1.5)

abline(v=1, lty=1, col="black",  lwd=1.5)

legend(1.5, 2.5, c("EU Density QUALITYe","EU Density COSTe", 
                   paste("EU Normal QUALITYe mean =",round(mean(df.sust.EU$QUALITYe),2)) , 
                   paste("EU Normal COSTe mean=",round(mean(df.sust.EU$COSTe),2)) ,
                   paste("Italy QUALITYe =",round(mean(df.sust.EU$QUALITYe[df.sust.EU$Country == ID.EU]),2)),
                   paste("Italy COSTe =",round(mean(df.sust.EU$COSTe[df.sust.EU$Country == ID.EU]),2))), 
       col=c("blue", "red", "blue", "red", "blue", "red"), 
       bty="n", 
       y.intersp=1.5, 
       lty=c(1,1,2,2,1,1), 
       cex = .75,
       lwd=c(1,1,1,1,1,1))

```





```{r}
library(gmodels)

for (i in 1:nrow(df.sust.EU)) {
        if (df.sust.EU$COSTe[i] > 1){
                df.sust.EU$Sustainability[i]="NS"
        }
        if (df.sust.EU$COSTe[i] <= 1){
                df.sust.EU$Sustainability[i]="S"
        }
        if (df.sust.EU$QUALITYe[i] >= 1){
                df.sust.EU$Quality[i]="Q"
        }
        if (df.sust.EU$QUALITYe[i] < 1){
                df.sust.EU$Quality[i]="NQ"
        }
}       

df.sust.EU

with(df.sust.EU, CrossTable(Sustainability, Quality, expected=T, digits=2, format = "SPSS"))

library(ReporteRs)
MyFTable <- vanilla.table(df.sust.EU)
MyFTable <- setFlexTableWidths( MyFTable, widths = c(3, 1, 0.5, rep(0.7, 6) ) )
MyFTable = setZebraStyle( MyFTable, odd = '#eeeeee', even = 'white' )
MyFTable
mydoc <- docx()
mydoc <- addFlexTable(mydoc, flextable = MyFTable)
writeDoc(mydoc, file = "MyFTable.docx")


```




# Write files to excel for export to ppt
```{r}
require(xlsx)
write.xlsx(df.EU.1990, "df.EU.xlsx", sheetName="1990")
write.xlsx(df.EU.2010, "df.EU.xlsx", sheetName="2010", append=TRUE)
write.xlsx(df.sust.EU, "df.EU.xlsx", sheetName="DELTA-EPSILON", append=TRUE)

```



